Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-12-10 manuscript/usses_water_ms.tex
more tiny revisions and matching up with response letter

\author[1,*]{Andrew T. Tredennick}

\corrauthor[1,*]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
\author[1,2,*]{Andrew R. Kleinhesselink}

@@

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\blfootnote{*Authors contributed equally to this work.}

@@

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita} and three perennial bunchgrasses,
\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
\emph{Hesperostipa comata}. During the period of our experiment (2011 --
2015), average mean annual precipitation was 265 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats, six of which were inside a large,
permanent livestock exclosure. We use these six plots as control plots
(i.e.~ambient precipitation) in the experiment described below.

@@

\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)}
from ambient conditions using a linear mixed effects model with
independent variables for treatment (irrigation or drought), season
(winter, spring, summer, fall), rainfall, and all two-way interactions.
Rainy days were defined as any day in which precipitation was recorded
and average temperature was above 3\(^{\circ}\)C. The day immediately
following rainfall was also classified as rainy. We fit the model using
the \texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
with random effects for plot group and date. We weighted observations by
the number of unique sensors or spot measurements that were taken in
each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for
which we took at least one ambient soil moisture measurement. This
procedure allowed us to predict daily soil moisture conditions for all
plots, even on those days when some of our direct treatment measurements
were missing due to malfunction of the sensors or dataloggers. See
\citet{Kleinhesselink2017b} for complete details on our approach to
estimating soil moisture.

@@

For each plot and year, we averaged the four readings for each
wavelength and then calculated a greenness index based on the same bands
used to calculate NDVI using the MODIS (Moderate Resolution Imaging
Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer)
bands for NDVI. We regressed the greenness index against the dry biomass
weight from the ten calibration plots to convert the greenness index to
ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
for each year and retained the regression with the better fit based on
\(R^2\) values. We then predicted ANPP using the best regression
equation for each year (Appendix 1).
\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}

@@

\noindent{}where \(\Sigma\) is the variance-covariance matrix and
\(j(i)\) reads as ``plot \(j\) associated with observation \(i\)''. The
random year effects (\(\boldsymbol{\eta}\)) are drawn from a normal
prior with mean 0 and standard deviation \(\sigma_{\text{year}}\), which
was drawn from a half-Cauchy distribution.
\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
A full description of our model and associated R \citep{R2016} code is
in Appendix 2.

@@

All R code and data necessary to reproduce our analysis has been
archived on Figshare (\emph{link here after acceptance}) and released on
GitHub (\url{https://github.com/atredennick/usses_water/releases/v0.1}).
We also include annotated Stan code in our model description in Appendix
2.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3B). The effect of soil moisture for each treatment is
associated with high uncertainty, however, with 95\% Bayesian credible
intervals that \new{broadly} overlap zero (Table 1). Although the
parameter estimates for the effect of soil moisture overlap zero, the
posterior distributions of the slopes all shrank and shifted to more
positive values relative to the prior distributions (Fig. A2-2), which
indicates the data did influence parameter estimates beyond the
information from the uninformative priors. Ecosystem functional response
was similar among treatments (Table 1; Fig. 3B), but there is evidence
that the \new{slopes for the drought and} irrigation treatments are less
than the slope for the control treatment. This evidence comes from
interpreting the posterior distributions of the slope offsets for the
treatments. \new{From these distributions,} we calculate a \new{91}\%
one-tailed probability that the estimate is less than zero for the
irrigation treatment
\new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment}
(Fig. 3A, right panel).

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. x). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. x)}.

Community composition was similar among treatments. The multidimensional
space of community composition overlapped among treatments in all years
and was equally dispersed in all years (Table 2; Fig. 4). Community
composition was also remarkably stable over time, with no evidence of
divergence among treatments (Table 2; Fig. 4).
\new{Splitting the analysis apart for perennials and annuals led to similar results.
The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
There is some evidence that the annual community changed in response to treatment in two years (Fig. Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}

@@

First, it could be that our precipitation manipulations were not large
enough to induce a response. That is, a 50\% decrease in any given year
may not be abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our treatments \emph{should} have been large
enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A4-1). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 4). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. For example, temperature can impact ANPP directly
\citep{Epstein1997} and by exacerbating the effects of soil moisture
\citep{DeBoeck2011}. Measurements of soil moisture likely contain a
signal of temperature, through its impact on evaporation and
infiltration, but the measurements will not capture the direct effect of
temperature on metabolic and physiological processes. We also did not
redistribute snow across our treatments in the winter, and snow melt may
spur early spring growth. Failure to account for potentially important
covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
3B) and the resulting uncertain relationship we observed between soil
moisture and ANPP across all treatments (Table 1, Fig. A2-2).
Unfortunately, we did not collect data on additional factors because our
focus was on precipitation.

@@

Another strategy to deal with variable environmental conditions is
avoidance, which would also result in a consistent ecosystem functional
response between drought and control treatments. For example, many of
the perennial grasses in our focal ecosystem avoid drought stress by
growing early in the growing season \citep[A.R. Kleinhesselink, personal
observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
in the soil profile thanks to a deep root system
\citep{Kulmatiski2017a}.
\new{The dominance of our site by the shrub \emph{A. tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.

\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
Ecosystem functional response may soon shift if sensitivity continues to increase.}

@@

Our results suggest the species in our focal plant community are
insensitive to to changes in precipitation regime,
\new{at least over the five years of our experiment}. Such insensitivity
can buffer species against precipitation variability in this semi-arid
ecosystem, making them successful in the long run. Longer, chronic
precipitation alteration might reveal plant community shifts that we did
not observe \citep[e.g.,][]{Wilcox2016}. For example, a long-term
increase in water availability could allow species that do not bet hedge
to gain prominence and dominate the ecosystem functional response.
\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
Our results suggest compositional shifts would have the largest impact
at high rainfall because the current community maintained consistent
ecosystem functional response at very low water availability.
\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
  2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.09 & 0.34 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.25 & 0.13 \\ 
  2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
  2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

>>>>

\author[1]{Andrew T. Tredennick}

\corrauthor[1]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
\author[1,2]{Andrew R. Kleinhesselink}

@@

@@

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita} and three perennial bunchgrasses,
\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
\emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2015), average mean annual precipitation was 265
mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats, six of which were inside a large,
permanent livestock exclosure. We use these six plots as control plots
(i.e.~ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}

@@

\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)}
from ambient conditions using a linear mixed effects model with
independent variables for treatment (irrigation or drought), season
(winter, spring, summer, fall), rainfall, and all two-way interactions.
Rainy days were defined as any day in which precipitation was recorded
and average temperature was above 3\(^{\circ}\)C. The day immediately
following rainfall was also classified as rainy. We fit the model using
the \texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
with random effects for plot group and date. We weighted observations by
the number of unique sensors or spot measurements that were taken in
each plot on that day.
\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
We could only predict soil moisture in the treated plots on days for
which we took at least one ambient soil moisture measurement. This
procedure allowed us to predict daily soil moisture conditions for all
plots, even on those days when some of our direct treatment measurements
were missing due to malfunction of the sensors or dataloggers. See
\citet{Kleinhesselink2017b} for complete details on our approach to
estimating soil moisture.

@@

For each plot and year, we averaged the four readings for each
wavelength and then calculated a greenness index based on the same bands
used to calculate NDVI using the MODIS (Moderate Resolution Imaging
Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer)
bands for NDVI. We regressed the greenness index against the dry biomass
weight from the ten calibration plots to convert the greenness index to
ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
for each year and retained the regression with the better fit based on
\(R^2\) values. We then predicted ANPP using the best regression
equation for each year (Appendix 2).
\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}

@@

\noindent{}where \(\Sigma\) is the variance-covariance matrix and
\(j(i)\) reads as ``plot \(j\) associated with observation \(i\)''. The
random year effects (\(\boldsymbol{\eta}\)) are drawn from a normal
prior with mean 0 and standard deviation \(\sigma_{\text{year}}\), which
was drawn from a half-Cauchy distribution.
\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
A full description of our model and associated R \citep{R2016} code is
in Appendix 3.

@@

All R code and data necessary to reproduce our analysis has been
archived on Figshare (\emph{link here after acceptance}) and released on
GitHub (\url{https://github.com/atredennick/usses_water/releases/v0.1}).
We also include annotated Stan code in our model description in Appendix
3.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3B). The effect of soil moisture for each treatment is
associated with high uncertainty, however, with 95\% Bayesian credible
intervals that \new{broadly} overlap zero (Table 1). Although the
parameter estimates for the effect of soil moisture overlap zero, the
posterior distributions of the slopes all shrank and shifted to more
positive values relative to the prior distributions (Fig. A3-2), which
indicates the data did influence parameter estimates beyond the
information from the uninformative priors. Ecosystem functional response
was similar among treatments (Table 1; Fig. 3B), but there is evidence
that the \new{slopes for the drought and} irrigation treatments are less
than the slope for the control treatment. This evidence comes from
interpreting the posterior distributions of the slope offsets for the
treatments. \new{From these distributions,} we calculate a \new{91}\%
one-tailed probability that the estimate is less than zero for the
irrigation treatment
\new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment}
(Fig. 3A, right panel).

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3D)}.

Community composition was similar among treatments. The multidimensional
space of community composition overlapped among treatments in all years
and was equally dispersed in all years (Table 2; Fig. 4). Community
composition was also remarkably stable over time, with no evidence of
divergence among treatments (Table 2; Fig. 4).
\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
Splitting the analysis apart for annuals led to similar results.
There is some evidence that the annual community changed in response to treatment in two years (Fig. A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig. A1-4).}

@@

First, it could be that our precipitation manipulations were not large
enough to induce a response. That is, a 50\% decrease in any given year
may not be abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our treatments \emph{should} have been large
enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A5-1). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 5). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. \new{For example,} \citet{LaPierre2016}
\new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
Moreover, temperature can impact ANPP directly \citep{Epstein1997} and
by exacerbating the effects of soil moisture \citep{DeBoeck2011}.
Measurements of soil moisture likely contain a signal of temperature,
through its impact on evaporation and infiltration, but the measurements
will not capture the direct effect of temperature on metabolic and
physiological processes. We also did not redistribute snow across our
treatments in the winter, and snow melt may spur early spring growth.
Failure to account for potentially important covariates could explain
the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
uncertain relationship we observed between soil moisture and ANPP across
all treatments (Table 1, Fig. A3-2). Unfortunately, we did not collect
data on additional factors because our focus was on precipitation.

@@

Another strategy to deal with variable environmental conditions is
avoidance, which would also result in a consistent ecosystem functional
response between drought and control treatments. For example, many of
the perennial grasses in our focal ecosystem avoid drought stress by
growing early in the growing season \citep[A.R. Kleinhesselink, personal
observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
in the soil profile thanks to a deep root system
\citep{Kulmatiski2017a}.
\new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.

\new{We did find that ANPP became more sensitive to drought over time (Fig. 3D).
Thus, the difference between ANPP, standardized by the difference in soil moisture, in drought and control plots grew over time.
The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
Ecosystem functional response may soon shift if sensitivity continues to increase.}

@@

Our results suggest the species in our focal plant community are
insensitive to to changes in precipitation regime,
\new{at least over the five years of our experiment}. Such insensitivity
can buffer species against precipitation variability in this semi-arid
ecosystem, making them successful in the long run. Longer, chronic
precipitation alteration might reveal plant community shifts that we did
not observe \citep[e.g.,][]{Wilcox2016},
\new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
Likewise, a long-term increase in water availability could allow species
that do not bet hedge to gain prominence and dominate the ecosystem
functional response.
\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
Our results suggest compositional shifts would have the largest impact
at high rainfall because the current community maintained consistent
ecosystem functional response at very low water availability.
\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
  2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.09 & 0.34 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.25 & 0.13 \\ 
  2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
  2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}
